Exercise 3: Software Architecture in Practice Case: Beer Web Store

### 2008-03-17 trunk/SAiP/Module1/H3/doc/qualityattributes.tex
Adding correctional reading

Finding out the requirements for the architecture is therefore more a matter of finding out what is important, more than it is to define what the product must do. The people who are able to define what is important to the architecture is naturally the stakeholders that the architecture affects, i.e. the owner, the user, the developer, the special interest groups, etc. Getting all of these peoples, often conflicting, requirements assembled and, even more importantly, prioritised, can be a difficult task. One way of doing so is to define scenarios, like the one mentioned above, and then having the stakeholders vote at the importance of the scenario.

@@

Once the scenarios are clarified and voted on, then the top priority scenarios may be formalized and it may be determined which quality attribute(s) they involve. For examples sake lets assume that the above scenario was selected as top priority.

@@

Naturally there are many more scenario required in order to properly cover all aspects of the quality properties.

@@

Finding out if a resource is down may be done using one of the fault detection tactics; ping/echo, heartbeat or exception. Due to the loose coupling between the client (browser) and the server, the exception solution is not beneficial. Since it may be assumed that the supervising process (remotely deployed) (the process monitoring if the server is down) is stable, then a heartbeat is sufficient to ensure availability, meaning that the web-server will issue a heart-beat every e.g. 1 second to a maintenance process on the same or another machine, and if one (or perhaps more) heartbeats is missed then the maintenance service can act.

The proper response to a faulty server may be many things, as defined by the tactics group \emph{fault recovery}. As the requirement for availability stipulated a three stage fault recovery, the server or servers (as it may be required to have multiple serves to do load-balancing, depending on how many users must be supported) can service the 5 second requirement, and then if an error occur (as detected by the heartbeat) then restart or swap to another server/database in 30 seconds. As 30 seconds is insufficient to boot a web server and a database, the Spare tactic would be insufficient. At the same time having Active redundancy would be over-kill, as a this kind of immediate switch-over is not required. For these reasons Passive Redundancy is sufficient.

As it is an internet system, it will also be sufficient to simply restart the session, and no \emph{Shadow}, \emph{State resynchronization} or \emph{Rollback} is needed in the main web server. The user may experience being thrown back to the main page, yet this is deemed acceptable, under availability and performance, yet it may not be in Usability. During a payment transaction, however, it must be ensured that the purchase is either completed or not-completed - halfway completed is not acceptable. for this it will be beneficial to use the \emph{Rollback} tactic on the \emph{Basket} component.

Finally a web-server may be deployed multiple times on teh same machine, allowing for the recovery or prevention of the lesser faults, by simply re-spawning the process that has become non-responsive, utilizing the \emph{Process Monitor} tactic. This is e.g. done by the Apache-server.

As it is not only the web server which may become non-responsive, it is also the database, it is also required to monitor the database. However, as a non-responsive data-base may be discovered by a non-responsive server, it is possible to use the ping-echo
 tactic with the database, and only perform the ping-echo if the server becomes non-responsive, to find out how much must be restarted (only the server or both the server and the database). 
Depending on the binding between the database and the server, it may also be possible to use the exceptions tactic between the server and the database, so the server itself can restart the database, if an exception occur in a call to the database.

@@

As we have all experience the irritation of a slow web-site, and as the competition on the internet is growing, it is important to achieve popper performance.

Due to the lowing price of hardware and the increasing price of development time, it is often cheapest and certainly simper to use the tactic \emph{Increase available resource}; introduce load balancing with multiple parallel servers thereby increase performance during heavy-load situations, ensure that the serves have sufficient network bandwidth, RAM and CPU strength to handle many simultanious database look-ups - here the tactic of \emph{Maintain multiple copies of data or computations} may be used to speed up database-look-up by caching the most popular searches, or cache the entire result in preparation for a search refinement. If the database is remotely located the caching of results becomes even more important.

>>>>

Finding out the requirements for the architecture is therefore more a matter of finding out what is important, more than it is to define what the product must do. The people who are able to define what is important to the architecture is naturally the stakeholders that the architecture affects, i.e. the owner, the user, the developer, the special interest groups, etc. Getting all of these peoples, often conflicting, requirements assembled and, even more importantly, prioritised, can be a difficult task. One way of doing so is to define scenarios, like the one mentioned above, and then having the stakeholders vote at the importance of the scenarios.

@@

Once the scenarios are clarified and voted on, then the top priority scenarios may be formalized and it may be determined which quality attribute(s) they involve. For examples lets assume that the above scenario was selected as top priority.

@@

Naturally there are many more scenario required in order to properly cover all aspects of the quality attributes.

@@

Finding out if a resource is down may be done using one of the fault detection tactics; ping/echo, heartbeat or exception. Due to the loose coupling between the client (browser) and the server, the exception solution is not beneficial. Since it may be assumed that the supervising process (remotely deployed) (the process monitoring if the server is down) is stable, then a heartbeat is sufficient to ensure availability, meaning that the web-server will issue a heart-beat every e.g. 1 second to a maintenance process, and if one (or perhaps more) heartbeats is missed then the maintenance service can act accordingly.

The proper response to a faulty server may be many things, as defined by the tactics group \emph{fault recovery}. As the requirement for availability stipulated a three stage fault recovery, the server or servers (as it may be required to have multiple serves to do load-balancing, depending on how many users must be supported) can service the 5 second requirement, and then if an error occur (as detected by the heartbeat), a restart or swap to another server/database can be performed in 30 seconds. As 30 seconds is insufficient to boot a web server and a database, the \emph{Spare} tactic would be insufficient. At the same time having \emph{Active redundancy} would be over-kill, as this kind of immediate switch-over is not required. For these reasons \emph{Passive Redundancy} is sufficient.

As it is an Internet system, it will also be sufficient to simply restart the session, and no \emph{Shadow}, \emph{State resynchronization} or \emph{Rollback} is needed in the main web server. The user may experience being thrown back to the main page, yet this is deemed acceptable, under availability and performance, yet it may not be in Usability. During a payment transaction, however, it must be ensured that the purchase is either completed or not-completed - halfway completed is not acceptable. For this it will be beneficial to use the \emph{Rollback} tactic on the \emph{Basket} component.

Finally a web-server may be deployed multiple times on the same machine, allowing for the recovery or prevention of the lesser faults, by simply re-spawning the process that has become non-responsive, utilizing the \emph{Process Monitor} tactic. This is e.g. done by the Apache-server.

As it is not only the web server which may become non-responsive, it is also required to monitor the database. However, as a non-responsive data-base may be discovered by a non-responsive server, it is possible to use the \emph{ping-echo} tactic with the database, and only perform the ping-echo if the server becomes non-responsive, to find out how much must be restarted (only the server or both the server and the database). 
Depending on the binding between the database and the server, it may also be possible to use the exceptions tactic between the server and the database, so the server itself can restart the database, if an exception occur in a call to the database.

@@

As we have all experience the irritation of a slow web-site, and as the competition on the internet is growing, it is important to achieve propper performance.

Due to the decreasing price of hardware and the increasing price of development time, it is often cheapest and certainly simpler, to use the tactic \emph{Increase available resource}; introduce load balancing with multiple parallel servers thereby increase performance during heavy-load situations, ensure that the serves have sufficient network bandwidth, RAM and CPU strength to handle many simultanious database look-ups - here the tactic of \emph{Maintain multiple copies of data or computations} may be used to speed up database-look-up by caching the most popular searches, or cache the entire result in preparation for a search refinement. If the database is remotely located the caching of results becomes even more important.
